Depression and suicidality as costly signals of need

### 2021-04-01 Paper.rmd
Edited methods and limitations

# Introduction
Since Durkheim’s pioneering effort to use quantitative social science to explain suicide (cite), anthropologists, psychologists, and sociologists have made varied attempts at explaining depression and suicide. However, despite this effort and substantial investment by funding agencies, rates of depression and suicide attempts have remained largely unchanged over the last 100 years (confirm and cite), and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to their high costs and the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).

@@

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and their 15-year-old daughter after they lost their house in a fire. During this time, the participant's daughter becomes jealous of her cousin before accusing her uncle of sexually assaulting her.

@@

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is described as trying to arrange a dowry for their older daughter so she can marry the man she loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. However, it is unclear if the second man is as unattractive as the older daughter says or if she does not value her younger sister enough, since any additional investment in the dowry will come at the expense of the younger daughter.

@@

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] (check this. he might have a paper saying it's not associated with depression as people tend to think) and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than.

@@

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population. Despite this limitation, it is more diverse than most convenience samples [@KennedyEtAl2018; @DworkinEtAl2016], with turkers in the U.S. mainly differing in being younger, more educated, lower income, and more female biased than the greater U.S. population [@RossEtAl2010].

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools, with there being some indication that it might outperform university samples in a few important areas. This is seen in that while both MTurk workers and university students engage in problematic behaviors that might lessen data quality to similar amounts [@NeckaEtAl2016], workers on MTurk have been shown to be more attentive than samples of university students [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from potential bot use and respondents faking their location to take surveys in a language they do not understand well, with the former concern being largely unfounded (at least for full on bots? do we mention human-assisted?)[@KennedyEtAl2018] and the risk of the latter able to be minimized through well designed attention checks (cite) .

@@

To ensure participants were paying attention, we used a variety of attention checks, which we clearly labeled in an attempt to avoid demand characteristics. The first was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. Capitalization did not matter and no word contained repeat vowels. Failure to enter the correct vowels in order resulted in the participants being filtered out of the study. The second followed the vignette and involved asking three questions about the story that should be easy to answer for anyone paying attention.

We also filtered out those who read the vignette faster than one-third the time it took MG to read it as fast as possible without concern for comprehension immediately after the vignette.

Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed the attention checks were paid $1 for their time. (mention time here)

@@

The largest source of limitations of this study is its lack of ecological validity, something which had to be sacrificed for any experimental design involving depression and suicide attempts. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction due to the lack of meaningful costs making support feel less costly or through the likely inability of participants to imagine themselves in an anonymous scenario (cite your paper on this) leading them to take an opportunity to signal generosity cheaply. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the bargaining of the costlier signals. However, both the findings of a main effect of signal and the effect of signal type on likelihood of support was largely mediated by change in belief suggest that participants were paying attention to the theoretically relevant details in ways that influenced their behavior.

Another limitation is our uncertainty over the data quality from the Indian sample. Compared to the US sample, far more participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses. Assuming this corresponded with more low-quality responses among those who passed the attention checks, the differences seen among those in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenarios or the effectiveness of the signals among Indians compared to Americans.

>>>>

# Introduction
Since Durkheim’s pioneering effort to use quantitative social science to explain suicide (cite), anthropologists, psychologists, and sociologists have made varied attempts at explaining depression and suicide. Despite this effort and substantial investment by funding agencies, rates of depression and suicide attempts have remained largely unchanged over the last 100 years (confirm and cite), and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to their high costs and the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).

@@

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and their 15-year-old daughter after their sister's family lost their house in a fire. During this time, the participant's daughter becomes jealous of her cousin before accusing her uncle of sexually assaulting her.

@@

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is described as trying to arrange a dowry for their older daughter so she can marry the man she loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Due to the younger daughter in the story only being two years younger, any additional investment in the dowry will come at the younger daughter's expsense.

@@

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] (check this. IIRC he might have a more recent paper saying it's not associated with depression as people tend to think) and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than. The full text for the signals are availible in the SI.

@@

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population (Thomas and Clifford 2017). Despite this limitation, it is more diverse than most convenience samples [@KennedyEtAl2018(2020 now); @DworkinEtAl2016; (Thomas and Clifford 2017)], with turkers in the US mainly differing in being younger, more educated, and lower income than the greater U.S. population [@RossEtAl2010]; Boas et al. 2018). Indian samples are also likely to be younger and more educated than the general population but differ from US samples in that they are less diverse, have higher incomes relative to other Indians, and are more likely to come from regions with good internet access (Boas et al. 2018)

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools[@NeckaEtAl2016, Robinson et al. 2019], with there being some indication that it might outperform university samples due to its previously mentioned diversity (Thomas and Clifford 2017) or the overall attentiveness of turkers [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from the threat of bot use or respondents faking their location to take surveys in a language they do not understand well, with there being little evidence of the former [@KennedyEtAl2018] and the risk of the latter able to be minimized through well designed attention checks, timed responses, or general study design (Aguinis et al. 2020; Huang et al. 2015; Kennedy et al. 2020; Thomas and Clifford 2017).

@@

To reduce the risk of low quality responses, we used a variety of attention checks, which we clearly labeled in an attempt to avoid demand characteristics. The first was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. Capitalization did not matter and no word contained repeat vowels. Failure to enter the correct vowels in order resulted in the participants being filtered out of the study. The second followed the vignette and involved asking three questions about the story that should be easy to answer for anyone paying attention.

We also filtered out those who read the vignette faster than one-third the time it took MG to read it as fast as possible without concern for comprehension. This occured immediately after participants finished the vignette.

Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed the attention checks were paid $1 for their time, with us estimating it would take 4-8 minutes before data collection. This study was certified exempt by the Washington State University Human Research Protection Program (certification number: 18528-001).

@@

The main limitation of this study is its lack of ecological validity, something which had to be sacrificed for any experimental design involving depression and suicide attempts. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction if the lack of meaningful costs made support feel less costly or if the likely inability of participants to imagine themselves in an anonymous scenario (cite your paper on this) led them to take an opportunity to signal generosity cheaply. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the costlier signals as bargaining tools. However, the fact that the relationships among signal cost, belief in need, and likelihood of providing support all largely fit existing theory (narrow and broad) indicates that participants paid attention to the relevant details and responded in ways that suggest some degree of external validity. For this reason, the benefits of addressing causal relationships that came through our experimental vignette approach likely outweighed the costs.

Another limitation is our uncertainty over the data quality from the Indian sample. Compared to the US sample, far more participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the differences seen among those in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenarios or the effectiveness of the signals among Indians compared to Americans.
